---
title: "Examples on using functions from ggplot2"
author: "Xiaona Zhou"
date: "4/17/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This R Markdown file contains sample codes for different types of plots that can be generate with `ggplot2` package. 

# Set up

Import packages

```{r global, include=FALSE}
library(ggplot2)
library(ggplot2movies)
library(dplyr)
library(hexbin)
library(ggthemes)
library(plotly)
```

# Histogram

1. Regular histogram

* `binwidth`: set the width of the bins 
* `bins`: set the number of bins in the histogram
* `alpha`: controls transparency. 0 means no fill color, 1 means solid  fill color
* `color`: color for the edges of the histogram
* `fill`: filling color in the histogram
```{r histogram, echo=TRUE, message=FALSE, warning=FALSE}
# create a histogram with binwidth 0.1, color the edge of the graph with red, color the inside of the histogram with pink, alpha = 0.5 means there's some transparency 
movies %>% 
ggplot(aes(x = rating)) +
  geom_histogram(binwidth = 0.1, color = 'red', fill = 'pink', alpha = 0.5 ) 
```

2. Better histogram

use count to color the histogram

```{r histogram_better, echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2movies)
library(dplyr)

movies %>% 
ggplot(aes(x = rating)) +
  geom_histogram(binwidth = 0.1, aes(fill = ..count..)) 
```



# Scatter plot

1. Basic Scatter plot 

* color can also be defined using hex color like '#33CCCC', check https://www.color-hex.com/
```{r scatter, echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg))+
  geom_point(size = 5, color = '#33CCCC')

```

2. Change features of the points with variables in Scatter plot 

Example 1.1: change the size of the points with variables in Scatter plot 
```{r scatter1, echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(aes(size = hp))
```

Example 1.2:
Note:use factor() because cyl is not continues. cyl contains only even numbers (4,6,8)
```{r scatter2, echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(aes(size = factor(cyl)))

```

Example 2.1: change the shape and color of the points with variables
```{r scatter3, echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(aes(shape = factor(cyl), color = factor(cyl)), size = 4)

```


Example 2.2: `scale_color_gradient` sets the colors for the points based on the value of `hp` 
```{r scatter4, echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(aes(color = hp),size = 5) + 
  scale_color_gradient(low = 'blue', high = 'red')

```


Example 3: `geom_smooth` adds a smooth line that fits the points
```{r scatter5, echo=TRUE, message=FALSE, warning=FALSE}
txhousing %>% 
  ggplot(aes(x = sales, y = volume)) +
  geom_point(aes(color = sales), alpha = 0.6) +
  scale_color_gradient(low = 'blue', high = 'red')+
  geom_smooth(color = 'green')

```



# Bar plot

Example 1:

```{r bar1, echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = class)) +
  geom_bar( color = 'blue', fill = 'blue')
```

Example 2: 
```{r}
mpg %>% 
  ggplot(aes(x = hwy)) +
  geom_bar(fill = '#F59D92')
```


Example 3: use `fill = drv` to add more information to the bar plot; 

```{r bar2, echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = class)) +
  geom_bar(aes(fill = drv))
```

Example 4: use `coord_flip` to switch x and y axis. 

```{r}

mpg %>% 
  ggplot(aes(x = manufacturer)) +
  geom_bar(aes(fill = factor(cyl))) +
  coord_flip()
```



Example 5: `position = 'dodge'` gives multiple bar plot side by side instead of stacking 

```{r bar3, echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = class)) +
  geom_bar(aes(fill = drv), position = 'dodge') 
```


Example 6: `position = 'fill'` shows the percentage of each class

```{r bar4, echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = class)) +
  geom_bar(aes(fill = drv), position = 'fill') 
```



# Boxplot

Example 1: `geom_bocplot` makes box plot

```{r boxplot1 , echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = factor(cyl), y = mpg)) +
  geom_boxplot()
```


Example 2:

```{r boxplot2 , echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = factor(cyl), y = mpg)) +
  geom_boxplot() +
  coord_flip()
```


Example 3: use `theme_dark()` to change the background

```{r boxplot3 , echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = factor(cyl), y = mpg)) +
  geom_boxplot(aes(fill = factor(cyl))) +
  theme_dark()
```



# Heat map 

Example 1: `geom_bin2d()` makes regular heat map 

```{r Variable_plotting1 , echo=TRUE, message=FALSE, warning=FALSE}
movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_bin2d()
```

Example 2:

```{r Variable_plotting2 , echo=TRUE, message=FALSE, warning=FALSE}
movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_bin2d() + scale_fill_gradient(high = 'red', low = 'green')
```

Example 3:

```{r Variable_plotting3 , echo=TRUE, message=FALSE, warning=FALSE}
movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_bin2d(binwidth = c(3,1)) + scale_fill_gradient(high = 'red', low = 'blue')
```


Example 4: `geom_hex()` makes heat map with hexagons 

```{r Variable_plotting4 , echo=TRUE, message=FALSE, warning=FALSE}
movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_hex() + scale_fill_gradient(high = 'red', low = 'blue')
```

# density plot

```{r Variable_plotting5 , echo=TRUE, message=FALSE, warning=FALSE}
movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_density2d() 

movies %>% 
  ggplot(aes(x = year, y = rating)) +
  geom_point()
```


# Coordanates Faceting

Example 1: use `coord_cartesian(xlim = c(1, 4), ylim = c(15,30))` to set the limits for x and y axis

```{r coor_facet1 , echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  coord_cartesian(xlim = c(1, 4), ylim = c(15,30))
```


Example 2: change the length of axis with `coord_fixed`

```{r coor_facet2 , echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  coord_fixed(ratio = 1/3)# ratio = x/y
```

Example 3: `facet_grid(.~cyl)` create multiple plots for different `cyl` values. `.~ cyl` split on x-axis

```{r coor_facet3 , echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(.~cyl)
```


Example 4: `facet_grid(drv~.)` create multiple plots for different `cyl` values. `drv~.` split on y-axis
Note:  `.~var_name` split on x-axis and `var_name~.` split on y-axis

```{r coor_facet4 , echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(drv~.)
```


Example 5: `facet_grid(y~x)` split on both y-axis and x-axis

```{r coor_facet5 , echo=TRUE, message=FALSE, warning=FALSE}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(drv~cyl)
```


# Themes

Example 1:

* theme_set(theme_minimal()) set the themes for all the plots 

```{r Themes1 , echo=TRUE, message=FALSE, warning=FALSE}
theme_set(theme_minimal())
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point()
```


Example 2: `theme_dark()`

```{r Themes2 , echo=TRUE, message=FALSE, warning=FALSE}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  theme_dark()
```



Example 3: load the library `ggthemes` and use `?ggthemes` to find out more themes


```{r Themes3 , echo=TRUE, message=FALSE, warning=FALSE}
library(ggthemes) 
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() + 
  theme_economist()
```

# Exercise

```{r as1, echo=TRUE, message=FALSE, warning=FALSE}
df <- read.csv('Economist_Assignment_Data.csv')
head(df)
df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region))


df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3)

df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3) +
  geom_smooth(aes(group=1))

df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3) +
  geom_smooth(aes(group = 1),method = 'lm', formula = y ~ log(x), se = FALSE, color = 'red')


df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3) +
  geom_smooth(aes(group=1),method = 'lm', formula = y ~ log(x),
              se = FALSE, color = 'red') +
  geom_text(aes(label = Country, color = Region)) 


pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan", "Afghanistan", "Congo", "Greece", "Argentina", "Brazil", "India", "Italy", "China", "South Africa", "Spane", "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France", "United States", "Germany", "Britain", "Barbados", "Norway", "Japan", "New Zealand", "Singapore")

df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3) +
  geom_smooth(method = 'lm', formula = y ~ log(x),
              se = FALSE, color = 'red') +
   geom_text(aes(label = Country), color = "gray20", 
                data = subset(df, Country %in% pointsToLabel),check_overlap = TRUE) +
  theme_bw() +
  scale_x_continuous(name = "Corruption Perceptions Index, 2011 (10=least corrupt)",limits = c(.9, 10.5),breaks=1:10) +
  scale_y_continuous(name = "Human Development Index, 2011 (1=Best)",limits = c(0.2, 1.0)) +
  ggtitle("Corruption and Human development") +
  theme_economist_white()

```


Label only a part of data 

1. Create a vector that contains all the labels to be showed on the plot, like pointsToLabel.
2. In geom_text(), use data = subset(data_name, variable_name %in% vector_name, check_overlap = TRUE)
```{r as2,echo=TRUE, message=FALSE, warning=FALSE}
pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan", "Afghanistan", "Congo", "Greece", "Argentina", "Brazil", "India", "Italy", "China", "South Africa", "Spane", "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France", "United States", "Germany", "Britain", "Barbados", "Norway", "Japan", "New Zealand", "Singapore")

df %>% 
  ggplot(aes(x = CPI, y = HDI)) +
  geom_point(aes(color = Region), shape = 'o', size = 3) +
  geom_smooth(method = 'lm', formula = y ~ log(x),
              se = FALSE, color = 'red') +
   geom_text(aes(label = Country), color = "gray20", 
                data = subset(df, Country %in% pointsToLabel),check_overlap = TRUE) +
  theme_bw() +
  scale_x_continuous(name = "Corruption Perceptions Index, 2011 (10=least corrupt)",limits = c(.9, 10.5),breaks=1:10) +
  scale_y_continuous(name = "Human Development Index, 2011 (1=Best)",limits = c(0.2, 1.0)) +
  ggtitle("Corruption and Human development") +
  theme_economist_white()

```


# Plotly

To use `plotly` to create interactive plots, simply add `ggplotly()` around the code. 

https://plotly.com/r/ 

```{r plotly1,echo=TRUE, message=FALSE, warning=FALSE}
ggplotly(mtcars %>%
  ggplot(aes(mpg ,wt)) +
  geom_point())
```

