---
title: "Examples on using functions from dyplr package"
author: "Xiaona Zhou"
date: "4/19/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
This R Markdown file contains examples on how to use various functions from `dyplr` package. I used `flights` data set from `nycflights13` package, and `mtcars` data set to demonstrate the usage of different functions. 

# Set up

Load `dplyr` and `nycflights13` packages. 
```{r global, warning=FALSE}
library(dplyr)
library(nycflights13)
```

Summary of the `flights`data set. 
```{r flight1, message=TRUE, warning=FALSE}
summary(flights)
```


Summary of the `mtcars` data set. 

```{r}
summary(mtcars)
```

# Examples

In the following section, you can found examples on how to use each functions with an brief explanation of the function, and the requirements for the inputs. 

## filter()
filter: select a subset of rows in a data frame.
filter(data, conditions separated by commas)

Example 1: select flights where `month` = 11, `day` = 3, and `carrier` name is 'AA'
```{r filter1, message=TRUE, warning=FALSE}
fl <- flights %>%
  filter(month == 11, day == 3, carrier == 'AA')
head(fl)
  
```


Example 2: select cars where `mpg` > 20 and `cyl` = 6
```{r filter2, message=TRUE, warning=FALSE}
mtcars %>%
  filter(mpg > 20, cyl == 6)
  
```


## slice()
slice: select rows by position 
slice(data, start_row_# : end_row_#)

Example 1: select the first 10 rows from `flights`
```{r slice1, message=TRUE, warning=FALSE}
flights %>% slice(1:10)
```


## arrange()

arrange: reorder the rows
arrange(data, year, month,day, arr_time): order by year, month, day, arr_time
Use desc(variable_name) if descending order is wanted 

Example 1: arrange the data by `cyl` in ascending order 

```{r arrange2, message=TRUE, warning=FALSE}
arr_1 <- mtcars %>% arrange(cyl)
head(arr_1)
```

Example 2: arrange the data first by `year`, and then by `month`, by `day`, lastly by `arr_time` (in descending order)
```{r arrange1, message=TRUE, warning=FALSE}
arr <- flights %>% arrange(year, month,day,desc(arr_time))
head(arr)
```


## select

select: select the columns wanted 
select(data, column_names)

Example 1: select columns `carrier` and `month`
```{r select1, message=TRUE, warning=FALSE}
sele <- flights %>% select(carrier, month,day)
head(sele)
```

Example 2: select columns `mpg`, `hp`

```{r select2, message=TRUE, warning=FALSE}
mtcars %>% select(mpg,hp)

```


## rename

rename: rename the columns 
rename(data, new_column_name = old_column_name)

Example 1: change column name `carrier` to `airline_carrier`. Note: the order is new column name = old column name, not the other way around. 
```{r rename1, message=TRUE, warning=FALSE}
rename_fl <- flights %>% rename(airline_carrier = carrier) %>% 
  select(airline_carrier)
head(rename_fl)
```

## distinct

distinct: select distinct values or unique values in a column
distinct(select(data, column_name))

Example 1: return all carrier names without repetition. 
```{r distinct1, message=TRUE, warning=FALSE}
distinct(select(flights, carrier))
# or 
flights %>% distinct(carrier)

```


## mutate & transmute

mutate: add new columns that are functions of existing columns
mutate(data, new_column_name = column_name - another_column_name)

transmute: similar to mutate but contains only the new column that has been created

Example 1: expected total travel time = schedule arrival time - schedule departure time. 

```{r mutate1, message=TRUE, warning=FALSE}
mutate_fl <- flights %>% 
  mutate(total_time = sched_arr_time - sched_dep_time) 
head(mutate_fl)
```

Same as above but with only the newly created column `total_time`
```{r}
transmute_fl <- flights%>% 
  transmute(total_time = sched_arr_time - sched_dep_time) 
head(transmute_fl)
```

Example 2: create a new column `Performance` by dividing `hp` by `wt`. 
```{r mutate2, message=TRUE, warning=FALSE}
mutate_car <- mtcars %>%
  mutate(Performance = hp/wt) 
head(mutate_car)
```

## summarise

summarise: collapse data frames into single rows using some sort of function that aggregates a result.
summarise(data, new_column_name = mean(column_name, na.rm = TRUE))

Example 1: calculate the mean `air_time` with `summarise` and `mean`. 

Note: avg_air_time contains the average air_time with rows contain NA removed (that's what 'na.rm = TRUE' do) 
```{r summarise1, message=TRUE, warning=FALSE}
flights %>% summarise(avg_air_time = mean(air_time, na.rm = TRUE))
```


Example 2: calculate the total `air_time` with `summarise` and `sum`

```{r}
flights %>% summarise(total_air_time = sum(air_time, na.rm = TRUE))
```

Example 3: calculate the average `mpg` across entire data set. 

```{r summarise2, message=TRUE, warning=FALSE}
mtcars %>% summarise(avg_mpg = mean(mpg)) #the mean mpg
```

Example 4: calculate the standard deviation pf hp value for cars with 6 cylinders.

```{r}
mtcars %>% filter(cyl == 6) %>% 
  summarise(std_hp = sd(hp))  #the standard deviation pf hp value for cars with 6 cylinders.

```



## sample_n & sample_frac

sample_n: select random number of rows
sample_n(data, # rows to select)


sample_frac: select a fraction of the rows
sample_frac(data, fraction of rows to select)

Example 1: randomly select 10 rows from the data frame
```{r sample1, message=TRUE, warning=FALSE}
sample_n(flights, 10)
```

Example 2: randomly select 10% of the rows from data frame

```{r}
sample_frac(flights,0.1)
```

To be continued...