---
title: "Examples on using functions from dyplr package"
author: "Xiaona Zhou"
date: "4/19/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This R Markdown file contains examples on how to use various functions from `tidyr` package. 

# Set up

Import packages

```{r global, message=FALSE, warning=FALSE}
library(tidyr)
library(tidyverse)
library(data.table)
```

Make two toy data set `df` and `stocks` for demonstration the functions

```{r}
comp <- c(1,1,1,2,2,2,3,3,3)
yr <- c(1998,1999,2000,1998,1999,2000,1998,1999,2000)
q1 <- runif(9, min=0, max=100)
q2 <- runif(9, min=0, max=100)
q3 <- runif(9, min=0, max=100)
q4 <- runif(9, min=0, max=100)

df <- data.frame(comp=comp,year=yr,Qtr1 = q1,Qtr2 = q2,Qtr3 = q3,Qtr4 = q4)
head(df)

stocks <- data.frame(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
head(stocks)
```

# Examples

In the following section, you can found examples on how to use each functions with an brief explanation of the function, and the requirements for the inputs.

## gather

gather: restructure the data 
gather(data, key, value, column_names)

Example 1: stack `Qrt1`, `Qrt2`,`Qrt3`,`Qrt4` into a new column `Quarter`.
```{r gather1, message=FALSE, warning=FALSE}
df_gathered <- df %>% 
  gather(Quarter, Revenue, Qtr1:Qtr4)

head(df_gathered)
```


Example 2: stack `X`, `Y`, `Z` into a new column `stock`
```{r gather2, message=FALSE, warning=FALSE}
stock_gathered <- stocks %>% 
  gather(stock, price, X:Z)
head(stock_gathered)
```

## spread

spread: undo gather()
spread(column_to_split_on, value)

Example 1: convert `stock_gathered` back to original format 
```{r spread, message=FALSE, warning=FALSE}

stock_gathered %>% spread(stock,price)# split different values within column 
#           stock into different columns, fill in corresponding price values. 
```

Example 2: spread the data set out base on time
```{r}
stock_gathered %>% spread(time, price)# split different values within column 
      #time into different columns, fill in corresponding price values. 
```


## separate

separate: turn a single character column into multiple columns
separate(data = data, col = target_column_name, into = c('new_col_name_1','new_col_name2'))
Note: by default, it is separated by non alphanumeric values, like '.' and '-'

Create a toy data set with only one column x
```{r separate1, message=FALSE, warning=FALSE}
df1 <- data.frame(x = c(NA, "a.x", "b.y", "c.z"))
df1
```

Example 1: separate the values in column `x` into two columns `ABC` and `XYZ`. when argument `sep` is not specified, it is separated by non alphanumeric values, like '.' and '-'. See example 2. 
```{r}
df1 %>% separate(col = x, into = c("ABC", "XYZ"))
```

Example 2: separate by '+'

```{r separate2, message=FALSE, warning=FALSE}
#create another toy data set
df2 <- data.frame(x= c('1+0', "1+0.5", "1+1", "1+2"))
df2
df2_se<- separate(data = df2, col=x, into = c("abc", "xyz"), sep = "([+])")# separate on +
df2_se
```


## unite

unite: undo separate; combine values in multiple columns into one 
unite(data = data, col_names_to_be_joined_separated_by_commas, sep = separator to use between joined values.)

Example 1: convert `df2_se` back to `df2` 
```{r unite, message=FALSE, warning=FALSE}
df2_se %>% 
  unite(new_joined, abc,xyz, sep = '+' )

```


## Join tables
1. inner_join(data1, data2) keeps the common rows and drops other rows
2. left_join(data1, data2) keeps all the rows in data1 add the extra columns from data2 if they are in common, put NA when no data available.
Note: right_join() is similar to left. Same result returns when switch data1 and data 2. 

3. full_join(data1, data2) keeps all rows from both data set, use NA when no data is available. 

4. anti_join: shows the rows in the left table that don't have a match in the right one.
Note: Most commonly you'd use it to check that rows weren't dropped from the left table.For instance, you could check that the number of rows in an anti join is zero.


```{r }
# create two tables with a common column name x1

df1 <- tibble(x1 = c('A','B','C','E','A'),
              x2 = c(1,2,3,4,5))
df2 <- tibble(x1 = c('A','B','D'),
              x3 = c(T,F,T))
df1
df2
inner_join(df1,df2)
left_join(df1,df2)
left_join(df2, df1)
full_join(df1, df2)
anti_join(df1,df2)#shows the rows that have been dropped by inner_join()
nrow(anti_join(df1,df2))==0
```


To be continued...